---
title: "R Notebook"
output: html_notebook
---

```{r}
# install packages(c("readxl", "dplyr", "lubridate"))
library(readxl)
library(dplyr)
library(lubridate) ## package to deal with dates
```

Read the data in

```{r}
beer <- read_excel(path = "../data-raw/beer-eng.xlsx")

head(beer)
```

Formating of dates

%Y -- YYYY (eg. 2010)
%m -- MM (eg. 11)
%d -- DD (eg. 02)

```{r}
## extract dates from the dataset by using as.Date function
beer <- beer %>% 
  mutate(date = as.character(date),
         date = as.Date(date, format = "%Y%m%d"))

### extract year from date variable using lubridate::year
beer <- beer %>% mutate(year = year(date) - 2009,
                        week_day = wday(date, week_start = 1))

head(beer)
```

Example usage of `lubridate::ymd` function


```{r}
ymd(c("2018-01-02", "2018.01.02", "2018/01/02"))
```

Check the size -- number of transactions (assuming that on given day there was only one transaction)

```{r}
dim(beer)
```

How many hh we have in our data and how many transactions they made

```{r}
beer %>% count(id, sort = TRUE) ## count ant sort
beer %>% count(id, sort = FALSE)  ## count without sorting
beer %>% count(id, sort = TRUE) %>% filter(n == 1) ## count and filter households with only one transaction
```

Let's see the distribution of count variable (number of beers bought)

```{r}
barplot(table(beer$count))
```

Let's built a simple linear regression

lm - linear model
formula - writting down the model

IN R: dependent ~ independent | y ~ x1 + x2 + x3 ... 
IN SAS/SPSS: y = x1 x2 x2 

$$
y_i  = \beta_0 + \epsilon_i, \quad \epsilon_i \sim N(0, \sigma)
$$

Fit the linear model

```{r}
model1 <- lm(formula = count ~ 1, data = beer)
model1
summary(model1)
```

Include into the model variable year

$$
y_i = \beta_0 + \beta_1 \times year + \epsilon_i, \quad \epsilon_i \sim N(0, \sigma)
$$
and include categorical variable

$$
y_i = \beta_0 + \beta_1 \times year + \beta_2 \times (wday=2) + \beta_3 * (wday=3) + ... + \epsilon_i, \quad \epsilon_i \sim N(0, \sigma)
$$

Fit the models in R

```{r}
model2 <- lm(formula = count ~ year, data = beer)
model3 <- lm(formula = count ~ year + factor(week_day), data = beer)
summary(model3)
```

numerical scale - number of beers, salary
nominal scale - region, week day 


Let's declare categorical variable in the dataset  

```{r}

beer <- beer %>% mutate(week_day = factor(week_day)) ## declare as factor / categorical variable
beer <- beer %>% mutate(week_day = relevel(week_day, ref = "5")) ## change the reference level

model3 <- lm(formula = count ~ year + week_day, data = beer)
summary(model3)
```

Standard asumptions of linear regression:

- linear relationship -- we don't know, we need to investigate
- normally distributed residuals N(0, sigma), constant variance (homoscedasticity) -- not met, standard errors will be underestimated
- continous Y variable -- count data -- linear regression is a bad choice, we should go for generalized linear models (GLM)
- observations are not correlated / independent from each other

