---
title: "R Notebook"
output: html_notebook
---

```{r}
install.packages("margins")
library(margins)
```

Assume the following model:

$$
Y_i  = \beta_0 + \beta_1 X_{1i} + \beta_2 X_{2i} + \beta_3 X_{1i} X_{2i}
$$

Margianl effect of $X_1$ would be

$$
\frac{\partial Y_i}{\partial X_{1i}} = \beta_1 + \beta_3 X_{2i}
$$

In general, we have three methods: 

+ Marginal effects at representative values (MERs), $X_{2i} = 2$

$$
MER = \frac{\partial Y_i}{\partial X_{1i}} = \beta_1 + \beta_3 X_{2i} = \beta_1 + \beta_3 * 2
$$

+ Marginal effect at means (MEMs), $X_{2i} = \bar{X}_{2i} = 25.34$

$$
MEM = \frac{\partial Y_i}{\partial X_{1i}} = \beta_1 + \beta_3 X_{2i} = \beta_1 + \beta_3 * \bar{X}_{2i} = \beta_1 + \beta_3 * 25.34
$$

+ Average marginal effect (AMEs)

$$
AME = \frac{\sum_i \frac{\partial Y_i}{\partial X_{1i}}}{n} = \frac{\sum_i \beta_1 + \beta_3 * X_{2i}}{n}
$$

```{r}
mtcars
```

Let's build a simple linear regression model based on mtcars dataset

```{r}
model1 <- lm(formula = mpg ~ cyl + hp + wt + am, data = mtcars)
summary(model1)
```

$$
mpg = 36.14  -0.75 * cyl - 0.03*hp - 2.61*wt + 1.48*am
$$

```{r}
model1_margins <- margins(model = model1)
summary(model1_margins)
```

We would like to get marginal effects for am = 1 or am = 0 or cyl = 6

```{r}
margins(model = model1, at = list(am = c(0, 1)))
```


```{r}
margins(model = model1, at = list(am = c(0, 1), cyl = 6))
```

We update model 1 by including interaction between am (automatic) and wt (weight)

X1 * X2 = X1 + X2 + X1:X2

```{r}
#model2 <- lm(formula = mpg ~ cyl + hp + wt + am + am*wt, data = mtcars)
model2 <- update(model1, . ~ . +  am*wt) ## . ~ . + am*wt 
summary(model2)
```

$$
Y_i  = \beta_0 + ... + \beta_3 WT + \beta_4 AM + \beta_5 WT*AM
$$

```{r}
model2_margins <- margins(model = model2)
summary(model1_margins)
summary(model2_margins)
```

Compare cars with am = 0 and am = 1

```{r}
model2_margins_at <- margins(model = model2, at = list(am = c(0,1)))
model2_margins_at
```

